---
title: "Web Scraping Wikipedia's data on Canadian Prime Ministers"
date: today
date-format: long
author: "Timothius Prajogi"
format: pdf
editor: visual
thanks: "Code and data supporting this analysis is available at: https://github.com/prajogt/$$$.git"
bibliography: ref.bib
---

```{r}
#| echo: false
#| message: false
#| warning: false

#### Workspace setup ####
library(janitor)
library(tidyverse)
library(knitr)

```

The data was scraped from Wikipedia's page on Canadian Prime Ministers [@citewiki].

These delay statistics were downloaded, cleaned, parsed, analyzed, and visualized using R [@citeR], a statistical programming language, with package support from `tidyverse` [@citetidyverse], a collection of libraries which included the following packages that were utilized: 

* `ggplot2` [@citeggplot] 
* `dplyr` [@citedplyr] 
* `readr` [@citereadr] 
* `tibble` [@citetibble]

Additionally for webscraping libraries xml2 [@citexml] and rvest [@citervest] were used.


```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-pms_data
#| tbl-cap: "A sample of Canadian Prime Minister's Lifespans"

# Read in cleaned data
canadian_pms <- 
  read_csv(
     file = "data/clean_canadian_pms.csv",
     show_col_types = FALSE
  )

canadian_pms |>
  head() |>
  kable(
    col.names = c("Prime Minister", "Birth Year", "Death Year", "Age at death")
  )
```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-pms
#| fig-cap: "The lifespans of Canadian Prime Ministers"


canadian_pms |>
  mutate(
    still_alive = if_else(is.na(death), "Yes", "No"),
    death = if_else(is.na(death), as.integer(2024), death)
  ) |>
  mutate(name = as_factor(name)) |>
  ggplot(
    aes(x = birth, xend = death, y = name, yend = name, color = still_alive)
    ) +
  geom_segment() +
  labs(
    x = "Year of birth", y = "Prime minister", color = "PM is currently alive"
    ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

# Findings

ASDLKASLKDm

# Process

# References

